Predicting trends in the quality of state-of-the-art neural networks without access to training or testing data

### 2020-07-02 submission/methods.tex
minor edits

Let us write the objective/optimization function (parameterized by $\mathbf{W}_{l}$s and $\mathbf{b}_{l}$s) for aE DNN with $L$ layers, activation functions $h_{l}(\cdot)$, and $N\times M$ weight matrices $\mathbf{W}_{l}$ and biases $\mathbf{b}_{l}$,~as \nred{the minimization of a general Loss Function $\mathcal{L}$ over  N (labeled) training data instances $(\mathbf{x}_{i},y_{i})$. For a typical supervised classification problem, one has
\begin{equation}
\underset{\mathbf{W}_{l},\mathbf{b}_{L}}{\text{argmin}}\;\sum_{i=1}^{N}\left(\mathcal{L}(E_{DNN}(\mathbf{x}_{i})-y_{i})\right)
\end{equation}
where }
\begin{equation}
E_{DNN}=h_{L}(\mathbf{W}_{L}\times h_{L-1}(\mathbf{W}_{L-1}\times h_{L-2}(\cdots)+\mathbf{b}_{L-1})+\mathbf{b}_{L}),
\label{eqn:dnn_energy}
\end{equation}
\nred{the loss function $\mathcal{L}()$ can take on a myriad of forms\cite{Janocha2017}, and $E_{DNN}$ is what we call the Energy Landscape.}

@@

Assume we are given several pretrained DNNs, e.g., as part of an architecture series.
The models have been trained and evaluated on labeled data $\{d_{i},y_{i}\}\in\mathcal{D}$, using standard techniques.  
The pretrained pytroch model files are publicly-available, and the test accuracies have been reported online.  
In this study, we do not have access to this data, and we have not trained any of the models ourselves,
nor have we re-evaluated the test accuracies.
We expect that most well-trained, production-quality models will employ one or more forms of regularization, such as Batch Normalization (BN), Dropout, etc., and many will also contain additional structure such as Skip Connections, etc. 
Here, we will ignore these details, and will focus only on the pretrained layer weight matrices $\mathbf{W}_{l}$.
Our methodological approach is thus similar to a statistical meta-analysis, common in biomedical research.

@@

The first two metrics are well-known in ML; the last two deserve special mention.
The empirical parameter $\alpha$ is the Power Law (PL) exponent that arises in the recently-developed HT-SR Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
The bottom line from HT-SR Theory is that smaller values of $\alpha$ should correspond to models with better correlation over multiple size scales and thus to better models; see Appendix~\ref{sxn:theory-review-appendix} for more details.
Operationally, $\alpha$ is determined by analyzing DNN weight matrices using the publicly-available \emph{WeightWatcher} tool~\cite{weightwatcher_package} to fit the Empirical Spectral Density (ESD) of $\mathbf{X}$, i.e., a histogram of the eigenvalues, call it $\rho(\lambda)$, to a truncated PL, 
\begin{equation}
\rho(\lambda)\sim\lambda^{\alpha},\;\;\lambda\le\lambda_{max}.
\end{equation}
%where $\lambda_{max}$ \emph{is} the largest eigenvalue of $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$.
Each of these quantities is defined for a given layer $\mathbf{W}$ matrix.

>>>>

Let us write the objective/optimization function (parameterized by $\mathbf{W}_{l}$s and $\mathbf{b}_{l}$s) for aE DNN with $L$ layers, activation functions $h_{l}(\cdot)$, and $N\times M$ weight matrices $\mathbf{W}_{l}$ and biases $\mathbf{b}_{l}$,~as \nred{the minimization of a general Loss Function $\mathcal{L}$ over  N training data instances and labels  $\{\mathbf{x}_{i},y_{i}\}\in\mathcal{D}$
For a typical supervised classification problem, one has
\begin{equation}
\underset{\mathbf{W}_{l},\mathbf{b}_{L}}{\text{argmin}}\;\sum_{i=1}^{N}\left(\mathcal{L}(E_{DNN}(\mathbf{x}_{i})-y_{i})\right)
\end{equation}
where the loss function $\mathcal{L}()$ can take on a myriad of forms\cite{Janocha2017}, and 
\begin{equation}
E_{DNN}=h_{L}(\mathbf{W}_{L}\times h_{L-1}(\mathbf{W}_{L-1}\times h_{L-2}(\cdots)+\mathbf{b}_{L-1})+\mathbf{b}_{L})
\label{eqn:dnn_energy}
\end{equation}
is what we call the Energy Landscape function.  Notice the $E_{DNN}$ does not explicitly depend on the data, but, rather,
maps data instances vectors $(\mathbf{x}_i)$ to predictions (i.e. labels).
Therefore, we can analyze $E_{DNN}$ absent of any training or test data. }

@@

Assume we are given several pretrained DNNs, e.g., as part of an architecture series.
The models have been trained and evaluated on the labeled data $\{\mathbf{x}_{i},y_{i}\}\in\mathcal{D}$, using standard techniques.  
The pretrained pytroch model files are publicly-available, and the test accuracies have been reported online.  
In this study, we do not have access to this data, and we have not trained any of the models ourselves. 
%nor have we re-evaluated the test accuracies.
We expect that most well-trained, production-quality models will employ one or more forms of regularization, such as Batch Normalization (BN), Dropout, etc., and many will also contain additional structure such as Skip Connections, etc. 
Here, we will ignore these details, and will focus only on the pretrained layer weight matrices $\mathbf{W}_{l}$.
Our methodological approach is thus similar to a statistical meta-analysis, common in biomedical research.

@@

The first two metrics are well-known in ML; the last two deserve special mention.
The empirical parameter $\alpha$ is the Power Law (PL) exponent that arises in the recently-developed HT-SR Theory~\cite{MM18_TR, MM19_HTSR_ICML, MM20_SDM}.
The bottom line from HT-SR Theory is that smaller values of $\alpha$ should correspond to models with better correlation over multiple size scales and thus to better models; see Appendix~\ref{sxn:theory-review-appendix} for more details.
Operationally, $\alpha$ is determined by analyzing DNN weight matrices using the publicly-available \emph{WeightWatcher} tool~\cite{weightwatcher_package} to fit the Empirical Spectral Density (ESD) of $\mathbf{X}$, i.e., a histogram of the eigenvalues, call it $\rho(\lambda)$, to a truncated PL, 
\begin{equation}
\rho(\lambda)\sim\lambda^{\alpha},\;\;\lambda\in[\lambda_{min},\lambda_{max}].
\end{equation}
Each of these quantities is defined for a given layer $\mathbf{W}$ matrix and the PL fitting method. \cite{Clauset}
\nred{Here, where $\lambda_{max}$ \emph{is} the largest eigenvalue of $\mathbf{X}=\mathbf{W}^{T}\mathbf{W}$, 
and $\lambda_{min}$ is selected automaticlly to yield the best PL fit (which minimizes the K-S distance). }
